Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC). }

### 2016-01-25 ijcar2016/body.tex
Implementation in cvc4 section.

\textbf{TODO:} Mention monotonicity analysis and zero sorts.

>>>>

\paragraph{Optimizations for Finding Finite Models of $\conv(\varphi)$}
Like other systems,
the finite model finding capability of \cvc incrementally fixes bounds on the cardinalities of uninterpreted types,
and increases these bounds until it encounters a model.
When multiple sorts are present, it uses a fairness scheme which
bounds the sum of cardinalities of all uninterpreted types~\cite{reynolds2013finite}.
For instance, if a signature has two uninterpreted types $\tau_1$ and $\tau_2$,
it will first search for models where 
the cardinality of $\tau_1$ plus the cardinality of $\tau_2$ is at most $2$,
then $3$, $4$, and so on.
To accelerate the search for models,
we use an optimization based on statically inferring \emph{monotone} types.
A monotone type is one in which models can always be extended with an additional element of that type.
Inference of monotone types has been show in previous work~\cite{claessen2011mono,korovin-2013}
to be an important component of finite model finding systems.
Moreover, it is easy to see that types $\fargtype{\con{f}}$ introduced by our translation $\conv$ are monotone.
The finite model finding capability of \cvc may take advantage of this fact by fixing the bounds for all 
monotone types simultaneously.
That is, if $\tau_1$ and $\tau_2$ are both monotone types (either introduced by our translation, or are otherwise inferred to be),
then we fix the bound for \emph{both} of these types to be $1$, then $2$, and so on.
This scheme allows the solver greater flexibility with respect to the default scheme,
and comes with no loss of generality with respect to models, since monotone sorts can always be extended to have equal cardinalities.

By default, 
\cvc uses techniques to minimize the number of literals it considers when constructing propositional satisfying assignments for formulas~\cite{relevancy2007}.
However, we have found such techniques degrade performance for finite model finding on
benchmarks having recursive functions that are defined by cases.
For this reason, we prefer disabling such techniques for benchmarks coming from our translation.

\ajr{More?}
